---
title: "Example `setupVerticalO2()` workflow"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Example setupVerticalO2 workflow}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5)
```

## Purpose

This vignette shows a minimal, serial `setupVerticalO2()` workflow using the
bundled CSV profile table at `data/profiles_sites_hist_vs_ssp585_long.csv`.
It loads one site/scenario profile, builds a FEISTY setup with oxygen and light
enabled, runs a short R simulation, and makes the most useful diagnostic plots.

The profile CSV is read outside `setupVerticalO2()`. The setup function receives
only the prepared profile data frame.

## Load FEISTY and one profile

```{r setup-profile}
devtools::load_all(path = here::here())
source(here::here("scripts", "vertical_o2_profiles.R"))

profile_path <- here::here("data", "profiles_sites_hist_vs_ssp585_long.csv")
stopifnot(file.exists(profile_path))

site <- "CCE"
scenario <- "hist"

profile <- read_vertical_o2_profile(
  profile_path = profile_path,
  site = site,
  scenario = scenario
)

head(profile)
summary(profile[, c("depth_mid_m", "temp_C", "pO2_kPa", "zmeso_day", "zmicro_day")])
```

## Plot the input forcing profiles

```{r forcing-profiles, fig.height=7}
op <- par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
ylim <- rev(range(profile$depth_mid_m))

plot(profile$temp_C, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "Temperature (deg C)", ylab = "Depth (m)", main = "Temperature", ylim = ylim)
plot(profile$pO2_kPa, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "pO2 (kPa)", ylab = "Depth (m)", main = "Oxygen", ylim = ylim)
plot(profile$zmicro_day, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmicro", ylab = "Depth (m)", main = "Microzooplankton", ylim = ylim)
lines(profile$zmicro_night, profile$depth_mid_m, lwd = 2, lty = 2)
plot(profile$zmeso_day, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmeso", ylab = "Depth (m)", main = "Mesozooplankton", ylim = ylim)
lines(profile$zmeso_night, profile$depth_mid_m, lwd = 2, lty = 2)
plot(profile$I_day_rel, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "relative light", ylab = "Depth (m)", main = "Light", ylim = ylim)
lines(profile$I_night_rel, profile$depth_mid_m, lwd = 2, lty = 2)
legend("bottomright", legend = c("day", "night"), lty = c(1, 2), lwd = 2, bty = "n")
plot(profile$dz_m, profile$depth_mid_m, type = "l", lwd = 2,
     xlab = "layer thickness (m)", ylab = "Depth (m)", main = "Vertical grid", ylim = ylim)
par(op)
```

## Build `setupVerticalO2()`

```{r build-setup}
p <- setupVerticalO2(
  profile = profile,
  nStages = 9,
  use_oxygen = TRUE,
  use_light = TRUE
)

c(
  setup = p$setup,
  n_stages = p$nStages,
  n_depth_layers = length(p$z_mid),
  min_glvl = min(p$glvl),
  max_glvl = max(p$glvl)
)
```

## Plot vertical habitat and limitation diagnostics

```{r setup-diagnostics, fig.height=7}
fish_group_names <- p$groupnames[p$nResources + seq_along(p$ix)]

op <- par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))
plot(p$phiLightDay, p$z_mid, type = "l", lwd = 2,
     xlab = "light multiplier", ylab = "Depth (m)", main = "Light used by setup",
     ylim = rev(range(p$z_mid)))
lines(p$phiLightNight, p$z_mid, lwd = 2, lty = 2)
legend("bottomright", legend = c("day", "night"), lty = c(1, 2), lwd = 2, bty = "n")

plot(p$mc[p$ixFish], p$glvl[p$ixFish], log = "x", pch = 16,
     xlab = "Body mass (g)", ylab = "oxygen limitation (glvl)",
     main = "Stage oxygen limitation", ylim = c(0, 1.05))
abline(h = c(0.25, 0.5, 0.75, 1), lty = 3, col = "grey80")

matplot(p$depthDay[, p$ixR[1:2]], p$z_mid, type = "l", lty = 1, lwd = 2,
        xlab = "depth fraction", ylab = "Depth (m)", main = "Resource depth: day",
        ylim = rev(range(p$z_mid)))
legend("bottomright", legend = p$groupnames[p$ixR[1:2]], col = 1:2, lty = 1, bty = "n")

matplot(p$depthNight[, p$ixR[1:2]], p$z_mid, type = "l", lty = 1, lwd = 2,
        xlab = "depth fraction", ylab = "Depth (m)", main = "Resource depth: night",
        ylim = rev(range(p$z_mid)))
legend("bottomright", legend = p$groupnames[p$ixR[1:2]], col = 1:2, lty = 1, bty = "n")
par(op)
```

## Run and plot a short simulation

```{r run-simulation}
sim <- simulateFEISTY(p = p, tEnd = 50, USEdll = FALSE)
```

```{r simulation-plots, fig.height=6}
plotSimulation(sim)
```

```{r network-and-rates, fig.height=5.5}
plotNetwork(sim)
plotRates(sim)
```

## Optional oxygen/light controls

For sensitivity checks, rerun the same profile with either switch disabled:

```{r controls, eval=FALSE}
p_no_oxygen <- setupVerticalO2(profile = profile, nStages = 9, use_oxygen = FALSE, use_light = TRUE)
p_no_light <- setupVerticalO2(profile = profile, nStages = 9, use_oxygen = TRUE, use_light = FALSE)
```
