---
title: "Parallel `setupVerticalO2()` site comparison maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel setupVerticalO2 site comparison maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 9, fig.height = 5)
```

## Purpose

This vignette runs `setupVerticalO2()` for any number of profile locations,
compares simulations **with oxygen limitation** against a **no-oxygen-limitation**
control (`glvl = 1` everywhere), saves compact biomass summaries, and maps the
results in a Robinson projection with black landmasses.

The workflow is designed for large profile CSV files:

1. Read the profile table once.
2. Split profiles by `(site, scenario)`.
3. Run each site in parallel.
4. Save only compact functional-group biomass summaries.
5. Plot total fish community biomass and oxygen-effect difference maps.

## User controls

Edit this chunk for your run. The key knobs are `n_cores` and
`run_first_n_locations` (set to `NULL` for all lon/lat cells in the NetCDF).

```{r controls}
devtools::load_all(path = here::here())

# --------------------------
# Obvious run controls
# --------------------------
# NetCDF source (hist only)
profile_nc_path <- here::here("data/profiles_hist_last10y.nc")

# Explicit parallel setting for server runs (edit as needed)
n_cores <- 32

# Start-up/test mode: run only first N locations to verify everything works.
# Set to NULL (or comment this line) for all available lon/lat locations.
run_first_n_locations <- 10

# Simulation controls
nStages <- 9
tEnd <- 50
tStep <- 1

out_dir <- here::here("outputs", "setupVerticalO2_parallel_maps")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```
```{r package-check}
map_pkgs <- c("sf", "rnaturalearth", "rnaturalearthdata")
missing_map_pkgs <- map_pkgs[!vapply(map_pkgs, requireNamespace, logical(1), quietly = TRUE)]
can_map <- length(missing_map_pkgs) == 0
if (!can_map) {
  message("Install these packages for maps: ", paste(missing_map_pkgs, collapse = ", "))
}
```

## Load and split the profile table once

For very large workflows, this chunk reads the NetCDF once and converts every
available lon/lat point into FEISTY long-profile rows (hist only).

```{r load-profiles}
stopifnot(file.exists(profile_nc_path))
if (!requireNamespace("ncdf4", quietly = TRUE)) {
  stop("Please install 'ncdf4' to read NetCDF profiles: install.packages('ncdf4')")
}

layer_bounds_from_midpoints <- function(mid_m) {
  mid_m <- as.numeric(mid_m)
  n <- length(mid_m)
  if (n <= 1) {
    top_m <- if (n == 0) numeric(0) else 0
    bot_m <- if (n == 0) numeric(0) else mid_m * 2
    dz_m <- bot_m - top_m
    return(list(top_m = top_m, bot_m = bot_m, dz_m = dz_m))
  }
  edges <- rep(NA_real_, n + 1)
  edges[2:n] <- 0.5 * (mid_m[1:(n - 1)] + mid_m[2:n])
  edges[1] <- max(0, mid_m[1] - 0.5 * (mid_m[2] - mid_m[1]))
  edges[n + 1] <- mid_m[n] + 0.5 * (mid_m[n] - mid_m[n - 1])
  list(top_m = edges[1:n], bot_m = edges[2:(n + 1)], dz_m = edges[2:(n + 1)] - edges[1:n])
}

nc <- ncdf4::nc_open(profile_nc_path)
on.exit(ncdf4::nc_close(nc), add = TRUE)

lon <- as.numeric(ncdf4::ncvar_get(nc, "lon"))
lat <- as.numeric(ncdf4::ncvar_get(nc, "lat"))
lev <- as.numeric(ncdf4::ncvar_get(nc, "lev"))

T_h <- ncdf4::ncvar_get(nc, "T_mean")
pO2_h <- ncdf4::ncvar_get(nc, "pO2_mean")
Z_h <- ncdf4::ncvar_get(nc, "zmeso_mean")
M_h <- ncdf4::ncvar_get(nc, "zmicro_mean")
Id_h <- ncdf4::ncvar_get(nc, "I_day")
In_h <- ncdf4::ncvar_get(nc, "I_night")

bnds <- layer_bounds_from_midpoints(lev)
depth_idx <- seq_along(lev)

build_profile_for_cell <- function(ix, iy) {
  lon360 <- lon[ix]
  lon180 <- ifelse(lon360 > 180, lon360 - 360, lon360)
  latg <- lat[iy]
  site_id <- sprintf("grid_%04d_%04d", ix, iy)

  data.frame(
    site = site_id,
    lon180_requested = lon180,
    lat_requested = latg,
    lon360_grid = lon360,
    lon180_grid = lon180,
    lat_grid = latg,
    depth_idx = depth_idx,
    depth_mid_m = lev,
    depth_top_m = bnds$top_m,
    depth_bot_m = bnds$bot_m,
    dz_m = bnds$dz_m,
    scenario = "hist",
    temp_C = as.numeric(T_h[ix, iy, ]),
    pO2_kPa = as.numeric(pO2_h[ix, iy, ]),
    zmeso = as.numeric(Z_h[ix, iy, ]),
    zmicro = as.numeric(M_h[ix, iy, ]),
    I_day_rel = as.numeric(Id_h[ix, iy, ]),
    I_night_rel = as.numeric(In_h[ix, iy, ]),
    stringsAsFactors = FALSE
  )
}

all_cells <- expand.grid(ix = seq_along(lon), iy = seq_along(lat))
if (!is.null(run_first_n_locations)) {
  all_cells <- utils::head(all_cells, run_first_n_locations)
}

profile_rows <- lapply(seq_len(nrow(all_cells)), function(i) build_profile_for_cell(all_cells$ix[i], all_cells$iy[i]))
profile_all <- do.call(rbind, profile_rows)

profile_all$key <- paste(profile_all$site, profile_all$scenario, sep = "::")
profile_list <- split(profile_all, profile_all$key)

# Persist converted profiles for reproducibility/debugging
utils::write.csv(profile_all, file.path(out_dir, "profiles_hist_from_nc_long.csv"), row.names = FALSE)

data.frame(
  n_locations = length(profile_list),
  min_lon = min(profile_all$lon180_grid, na.rm = TRUE),
  max_lon = max(profile_all$lon180_grid, na.rm = TRUE),
  min_lat = min(profile_all$lat_grid, na.rm = TRUE),
  max_lat = max(profile_all$lat_grid, na.rm = TRUE)
)
```

## Worker function

Each worker runs two matched simulations for one profile:

- `use_oxygen = TRUE`: profile oxygen affects `glvl`.
- `use_oxygen = FALSE`: no oxygen limitation (`glvl = 1` everywhere), while the
  same physical/biological profile is otherwise retained.

The saved output is intentionally compact: final functional-group biomasses and
total fish community biomass.

```{r worker}
summarise_final_biomass <- function(sim, condition) {
  p <- sim$p

  # Primary source: grouped biomass diagnostics from simulateFEISTY()
  final_group <- NULL
  if (!is.null(sim$totBiomass) && NROW(sim$totBiomass) >= sim$nTime && NCOL(sim$totBiomass) > 0) {
    final_group <- as.numeric(sim$totBiomass[sim$nTime, ])
    names(final_group) <- p$groupnames[-seq_len(p$nResources)]
  }

  # Fallback: reconstruct group totals from stage biomasses
  if (is.null(final_group) || length(final_group) == 0) {
    if (is.null(sim$B) || NROW(sim$B) < sim$nTime) {
      stop("No grouped biomass diagnostics and no usable stage biomass matrix in simulation output.")
    }

    final_group <- vapply(seq_along(p$ix), function(i) {
      cols <- p$ix[[i]] - p$nResources
      sum(as.numeric(sim$B[sim$nTime, cols]), na.rm = TRUE)
    }, numeric(1))
    names(final_group) <- p$groupnames[p$nResources + seq_along(p$ix)]
  }

  out <- data.frame(
    condition = condition,
    group = names(final_group),
    biomass = as.numeric(final_group),
    stringsAsFactors = FALSE
  )

  if (nrow(out) == 0) {
    stop("No rows in summarised biomass output for condition: ", condition)
  }

  out
}

run_one_location <- function(profile_one, key) {
  profile_one <- profile_one[order(profile_one$depth_idx, profile_one$depth_mid_m), ]
  profile_one$key <- NULL

  coord <- profile_one[stats::complete.cases(profile_one[, c("lon180_grid", "lat_grid")]), ][1, ]
  site <- as.character(profile_one$site[1])
  scenario <- as.character(profile_one$scenario[1])

  build_p <- function(use_oxygen) {
    setupVerticalO2(
      profile = profile_one,
      nStages = nStages,
      use_oxygen = use_oxygen,
      K_O = 2,
      h_O = 1,
      delta_pO2_ref = 2,
      Q10_D = 2,
      Q10_S = 1.5,
      K_L = 0.1,
      L_min = 0.5,
      L_max = 1.5,
      light_bottom_to_one = TRUE
    )
  }

  p_o2 <- build_p(TRUE)
  sim_o2 <- simulateFEISTY(p = p_o2, tEnd = tEnd, tStep = tStep, USEdll = FALSE)

  p_no_o2 <- build_p(FALSE)
  sim_no_o2 <- simulateFEISTY(p = p_no_o2, tEnd = tEnd, tStep = tStep, USEdll = FALSE)

  biomass <- rbind(
    summarise_final_biomass(sim_o2, "oxygen_profile"),
    summarise_final_biomass(sim_no_o2, "no_oxygen_limitation")
  )

  biomass$site <- site
  biomass$scenario <- scenario
  biomass$lon <- coord$lon180_grid
  biomass$lat <- coord$lat_grid
  biomass$key <- key

  total <- aggregate(biomass ~ condition + site + scenario + lon + lat + key, biomass, sum)
  total$group <- "total_fish"
  biomass <- rbind(biomass, total[, names(biomass)])

  wide <- reshape(
    biomass[, c("site", "scenario", "lon", "lat", "key", "group", "condition", "biomass")],
    idvar = c("site", "scenario", "lon", "lat", "key", "group"),
    timevar = "condition",
    direction = "wide"
  )
  names(wide) <- sub("^biomass\\.", "biomass_", names(wide))
  wide$biomass_difference <- wide$biomass_oxygen_profile - wide$biomass_no_oxygen_limitation
  wide$biomass_ratio <- wide$biomass_oxygen_profile / wide$biomass_no_oxygen_limitation

  list(biomass = biomass, difference = wide)
}
```

## Run sites in parallel and save outputs

The code uses base R `parallel` so no extra parallel package is required. On
Unix-like systems it uses forking; on Windows it uses a PSOCK cluster.

```{r run-parallel}
keys <- names(profile_list)

safe_run <- function(k) {
  tryCatch(
    run_one_location(profile_list[[k]], k),
    error = function(e) list(error = conditionMessage(e), key = k)
  )
}

if (.Platform$OS.type == "windows" && n_cores > 1) {
  cl <- parallel::makeCluster(n_cores)
  on.exit(parallel::stopCluster(cl), add = TRUE)
  parallel::clusterExport(
    cl,
    varlist = c("profile_list", "run_one_location", "summarise_final_biomass", "nStages", "tEnd", "tStep", "safe_run"),
    envir = environment()
  )
  parallel::clusterEvalQ(cl, devtools::load_all(path = here::here()))
  results <- parallel::parLapply(cl, keys, safe_run)
} else {
  results <- parallel::mclapply(
    keys,
    safe_run,
    mc.cores = min(n_cores, length(keys))
  )
}

ok <- vapply(results, function(x) is.list(x) && all(c("biomass", "difference") %in% names(x)), logical(1))
if (!all(ok)) {
  failed <- results[!ok]
  fail_tbl <- data.frame(
    key = vapply(failed, function(x) if (!is.null(x$key)) x$key else NA_character_, character(1)),
    error = vapply(failed, function(x) if (!is.null(x$error)) x$error else "Unknown error", character(1)),
    stringsAsFactors = FALSE
  )
  utils::write.csv(fail_tbl, file.path(out_dir, "failed_locations.csv"), row.names = FALSE)
  message("Failed locations: ", nrow(fail_tbl), ". See ", file.path(out_dir, "failed_locations.csv"))
}

results_ok <- results[ok]
if (length(results_ok) == 0) stop("All locations failed. See failed_locations.csv for details.")

biomass_summary <- do.call(rbind, lapply(results_ok, `[[`, "biomass"))
biomass_difference <- do.call(rbind, lapply(results_ok, `[[`, "difference"))

utils::write.csv(biomass_summary, file.path(out_dir, "functional_group_biomass_summary.csv"), row.names = FALSE)
utils::write.csv(biomass_difference, file.path(out_dir, "functional_group_biomass_differences.csv"), row.names = FALSE)
saveRDS(results, file.path(out_dir, "setupVerticalO2_parallel_results.rds"))

head(biomass_difference)
```

## Map helpers

```{r map-helpers, eval=can_map}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
robin_crs <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

plot_points_robin <- function(data, value, title, legend_title, palette = ggplot2::scale_color_viridis_c()) {
  pts <- sf::st_as_sf(data, coords = c("lon", "lat"), crs = 4326, remove = FALSE)
  ggplot2::ggplot() +
    ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1) +
    ggplot2::geom_sf(data = pts, ggplot2::aes(color = .data[[value]]), size = 2.5) +
    palette +
    ggplot2::coord_sf(crs = robin_crs) +
    ggplot2::labs(title = title, color = legend_title) +
    ggplot2::theme_minimal() +
    ggplot2::theme(panel.grid.major = ggplot2::element_line(color = "grey85", linewidth = 0.2))
}
```

## Total fish community biomass maps

```{r total-biomass-map, fig.height=5.5, eval=can_map}
total_biomass <- subset(biomass_summary, group == "total_fish")
p_total_biomass <- plot_points_robin(
  subset(total_biomass, condition == "oxygen_profile"),
  value = "biomass",
  title = "Total fish biomass with profile oxygen",
  legend_title = "g m^-2"
)
ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_oxygen_profile.png"), p_total_biomass, width = 10, height = 6, dpi = 300)
p_total_biomass
```

```{r total-difference-map, fig.height=5.5, eval=can_map}
total_diff <- subset(biomass_difference, group == "total_fish")
p_total_diff <- plot_points_robin(
  total_diff,
  value = "biomass_difference",
  title = "Total fish biomass difference: oxygen profile - no oxygen limitation",
  legend_title = "difference (g m^-2)",
  palette = ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)
)
ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_difference.png"), p_total_diff, width = 10, height = 6, dpi = 300)
p_total_diff
```

## Functional-group biomass difference maps

```{r group-difference-maps, fig.height=8, eval=can_map}
group_diff <- subset(biomass_difference, group != "total_fish")
group_diff_sf <- sf::st_as_sf(group_diff, coords = c("lon", "lat"), crs = 4326, remove = FALSE)

p_group_diff <- ggplot2::ggplot() +
  ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1) +
  ggplot2::geom_sf(data = group_diff_sf, ggplot2::aes(color = biomass_difference), size = 2.2) +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  ggplot2::coord_sf(crs = robin_crs) +
  ggplot2::facet_wrap(~ group) +
  ggplot2::labs(
    title = "Functional group biomass difference: oxygen profile - no oxygen limitation",
    color = "difference (g m^-2)"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(panel.grid.major = ggplot2::element_line(color = "grey85", linewidth = 0.2))

ggplot2::ggsave(file.path(out_dir, "map_functional_group_biomass_differences.png"), p_group_diff, width = 12, height = 8, dpi = 300)
p_group_diff
```

## Notes

- For full global runs from NetCDF, set `run_first_n_locations <- NULL` and keep `n_cores <- 32` (or edit).
- Outputs are saved in `out_dir`, so the full simulation objects do not need to
  be kept in memory after mapping.
- The no-oxygen control is intentionally `use_oxygen = FALSE`; do not replace
  the observed oxygen profile with an artificial high-pO2 column for this
  comparison.
