---
title: "Testing `setupVerticalO2()` from a Single NetCDF Coordinate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing setupVerticalO2 from NetCDF coordinate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5)
```

## Purpose

This test vignette is a runnable workflow for validating that a **single profile can be read directly from the NetCDF source** and then passed into `setupVerticalO2()`.

Default coordinate is the same CCE location used in the CSV workflow:

- longitude (180 convention): `-123`
- latitude: `34`

## Load FEISTY functions and configure NetCDF input

```{r setup-local}
# Load FEISTY functions locally through devtools.
devtools::load_all(path = here::here())

nc_hist_path <- here::here("data/profiles_hist_last10y.nc")
stopifnot(file.exists(nc_hist_path))

# Default coordinate matches the CCE row used in CSV-based tests.
site_name <- "CCE"
lon180_requested <- -123
lat_requested <- 34
scenario <- "hist"
```

## Read a single coordinate profile from NetCDF

```{r profile-from-nc}
stopifnot(requireNamespace("ncdf4", quietly = TRUE))

nc <- ncdf4::nc_open(nc_hist_path)
on.exit(ncdf4::nc_close(nc), add = TRUE)

lon <- as.numeric(ncdf4::ncvar_get(nc, "lon"))
lat <- as.numeric(ncdf4::ncvar_get(nc, "lat"))
lev <- as.numeric(ncdf4::ncvar_get(nc, "lev"))

lon360_requested <- ifelse(lon180_requested < 0, lon180_requested + 360, lon180_requested)
ix <- which.min(abs(lon - lon360_requested))
iy <- which.min(abs(lat - lat_requested))

lon360_grid <- lon[ix]
lon180_grid <- ifelse(lon360_grid > 180, lon360_grid - 360, lon360_grid)
lat_grid <- lat[iy]

cat("Requested (lon180, lat):", lon180_requested, lat_requested, "\n")
cat("Selected  (lon180, lat):", lon180_grid, lat_grid, "\n")
cat("Grid indices (ix, iy):", ix, iy, "\n")

temp_C <- as.numeric(read_var_lon_lat_lev(nc, "T_mean", length(lon), length(lat), length(lev))[ix, iy, ])
pO2_kPa <- as.numeric(read_var_lon_lat_lev(nc, "pO2_mean", length(lon), length(lat), length(lev))[ix, iy, ])
zmeso <- as.numeric(read_var_lon_lat_lev(nc, "zmeso_mean", length(lon), length(lat), length(lev))[ix, iy, ])
zmicro <- as.numeric(read_var_lon_lat_lev(nc, "zmicro_mean", length(lon), length(lat), length(lev))[ix, iy, ])
I_day_rel <- as.numeric(read_var_lon_lat_lev(nc, "I_day", length(lon), length(lat), length(lev))[ix, iy, ])
I_night_rel <- as.numeric(read_var_lon_lat_lev(nc, "I_night", length(lon), length(lat), length(lev))[ix, iy, ])

bnds <- layer_bounds_from_midpoints(lev)
profile_nc <- data.frame(
  site = site_name,
  lon180_requested = lon180_requested,
  lat_requested = lat_requested,
  lon360_grid = lon360_grid,
  lon180_grid = lon180_grid,
  lat_grid = lat_grid,
  depth_idx = seq_along(lev),
  depth_mid_m = lev,
  depth_top_m = bnds$top_m,
  depth_bot_m = bnds$bot_m,
  dz_m = bnds$dz_m,
  scenario = scenario,
  temp_C = temp_C,
  pO2_kPa = pO2_kPa,
  zmeso = zmeso,
  zmicro = zmicro,
  I_day_rel = I_day_rel,
  I_night_rel = I_night_rel,
  stringsAsFactors = FALSE
)

# Normalize with the same trimming/interpolation logic used by
# read_vertical_o2_profile(...), so profile= and profile_path= behave alike.
profile_nc <- normalize_vertical_o2_profile(
  profile_nc,
  fill_internal_gaps = FALSE
)

str(profile_nc)
head(profile_nc)
summary(profile_nc$dz_m)
```

## Plot vertical forcing profiles

```{r forcing-profiles, fig.height=8}
op <- par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
plot(profile_nc$temp_C, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "Temperature (C)", ylab = "Depth (m)", main = "Temperature",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$pO2_kPa, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "pO2 (kPa)", ylab = "Depth (m)", main = "Oxygen",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$zmicro, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmicro", ylab = "Depth (m)", main = "Microzoo",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$zmeso, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmeso", ylab = "Depth (m)", main = "Mesozoo",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$I_day_rel, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_day_rel", ylab = "Depth (m)", main = "Light (day)",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$I_night_rel, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_night_rel", ylab = "Depth (m)", main = "Light (night)",
     ylim = rev(range(profile_nc$depth_mid_m)))
par(op)
```

## Build `setupVerticalO2()` from NetCDF-derived profile object

```{r build-setup}
p <- setupVerticalO2(
  profile = profile_nc,
  nStages = 9,
  K_O = 2,
  h_O = 1,
  delta_pO2_ref = 2,
  Q10_D = 2,
  Q10_S = 1.5,
  K_L = 0.1,
  L_min = 0.5,
  L_max = 1.5,
  light_bottom_to_one = TRUE
)

# Compare to profile_path path behavior: setupVerticalO2() only reads from
# profile_path when profile is NULL.
cat("nrow(profile_nc):", nrow(profile_nc), "\n")
cat("depth range:", min(profile_nc$depth_mid_m), "to", max(profile_nc$depth_mid_m), "m\n")

p$setup
dim(p$depthDay)
dim(p$depthNight)
range(p$glvl)
range(p$phiLightDay)
range(p$phiLightNight)
```

## Run and inspect a short simulation

```{r run-sim}
sim <- simulateFEISTY(p = p, tEnd = 50, USEdll = FALSE)
```

## Proposal figure set: oxygen vs no-oxygen (CCE)

```{r proposal-setup, message=FALSE, warning=FALSE}
stopifnot(requireNamespace("ggplot2", quietly = TRUE))
stopifnot(requireNamespace("dplyr", quietly = TRUE))
stopifnot(requireNamespace("tidyr", quietly = TRUE))
stopifnot(requireNamespace("patchwork", quietly = TRUE))

proposal_theme <- function(base_size = 9) {
  ggplot2::theme_minimal(base_size = base_size) +
    ggplot2::theme(
      panel.grid.minor = ggplot2::element_blank(),
      panel.border = ggplot2::element_rect(colour = "grey50", fill = NA, linewidth = 0.4),
      axis.ticks = ggplot2::element_line(colour = "grey50"),
      axis.text = ggplot2::element_text(size = 7, colour = "grey20"),
      axis.title = ggplot2::element_text(size = 8),
      legend.position = "bottom",
      legend.title = ggplot2::element_text(size = 8),
      legend.text = ggplot2::element_text(size = 7),
      plot.title = ggplot2::element_text(size = 9, face = "bold")
    )
}

# Build the no-oxygen setup/simulation with the same CCE profile
p_no_o2 <- setupVerticalO2(
  profile = profile_nc,
  nStages = 9,
  use_oxygen = FALSE,
  K_O = 2,
  h_O = 1,
  delta_pO2_ref = 2,
  Q10_D = 2,
  Q10_S = 1.5,
  K_L = 0.1,
  L_min = 0.5,
  L_max = 1.5,
  light_bottom_to_one = TRUE
)
sim_no_o2 <- simulateFEISTY(p = p_no_o2, tEnd = 50, USEdll = FALSE)
```

### 1) Combined temperature + oxygen profile

```{r proposal-forcing-profile, fig.height=4.8}
forcing_long <- dplyr::bind_rows(
  data.frame(var = "Temperature", value = profile_nc$temp_C, depth_mid_m = profile_nc$depth_mid_m),
  data.frame(var = "Oxygen", value = profile_nc$pO2_kPa, depth_mid_m = profile_nc$depth_mid_m)
)

temp_rng <- range(profile_nc$temp_C, na.rm = TRUE)
o2_rng <- range(profile_nc$pO2_kPa, na.rm = TRUE)
rescale_to_temp <- function(x) (x - o2_rng[1]) / diff(o2_rng) * diff(temp_rng) + temp_rng[1]
unscale_from_temp <- function(x) (x - temp_rng[1]) / diff(temp_rng) * diff(o2_rng) + o2_rng[1]

forcing_plot <- ggplot2::ggplot() +
  ggplot2::geom_path(data = profile_nc, ggplot2::aes(x = temp_C, y = depth_mid_m),
                     linewidth = 1.0, color = "#BA584D", na.rm = TRUE) +
  ggplot2::geom_path(data = profile_nc, ggplot2::aes(x = pO2_kPa, y = depth_mid_m),
                     linewidth = 1.0, color = "#6a51a3", na.rm = TRUE) +
  ggplot2::scale_x_continuous(limits = c(0, 25), breaks = c(0, 5, 10, 15, 20, 25), name = NULL) +
  ggplot2::scale_y_reverse(expand = c(0.01, 0.01)) +
  ggplot2::labs(y = "Depth (m)") +
  proposal_theme() +
  ggplot2::theme(
    axis.title.x = ggplot2::element_blank(),
    aspect.ratio = 1.55
  )

forcing_plot
```

### 2) Shared forcing + fish profiles (matched depth axis)

```{r proposal-depth-distributions, fig.height=7.5}
find_group_col <- function(p, pattern) {
  idx <- grep(pattern, p$groupnames[-seq_len(p$nResources)], ignore.case = TRUE)
  if (length(idx) == 0) stop("Could not find group pattern: ", pattern)
  p$ix[[idx[1]]][length(p$ix[[idx[1]]])]
}

stage_cols <- data.frame(
  species = c("Large pelagic", "Mesopelagic", "Midwater predator"),
  col_id = c(
    find_group_col(p, "largePel"),
    find_group_col(p, "mesoPel"),
    find_group_col(p, "midwPred")
  )
)

# Use the same depth support as profile forcing when possible
z_depth <- if (!is.null(p$z_mid) && length(p$z_mid) == nrow(p$depthDay)) {
  p$z_mid
} else if (length(profile_nc$depth_mid_m) == nrow(p$depthDay)) {
  profile_nc$depth_mid_m
} else if (length(p$z) == nrow(p$depthDay)) {
  p$z
} else {
  seq_len(nrow(p$depthDay))
}

build_depth_df <- function(period = c("Day", "Night")) {
  period <- match.arg(period)
  mat <- if (period == "Day") p$depthDay else p$depthNight
  do.call(rbind, lapply(seq_len(nrow(stage_cols)), function(i) {
    ci <- stage_cols$col_id[i]
    sp <- stage_cols$species[i]
    data.frame(
      species = rep(sp, length(z_depth)),
      period = rep(period, length(z_depth)),
      depth_mid_m = z_depth,
      probability = as.numeric(mat[, ci]),
      stringsAsFactors = FALSE
    )
  }))
}

depth_dist <- dplyr::bind_rows(build_depth_df("Day"), build_depth_df("Night"))

species_order <- c("Large pelagic", "Mesopelagic", "Midwater predator")
y_lim <- range(z_depth, na.rm = TRUE)

mk_species_panel <- function(sp) {
  ggplot2::ggplot(subset(depth_dist, species == sp), ggplot2::aes(x = probability, y = depth_mid_m, linetype = period)) +
    ggplot2::geom_path(linewidth = 0.9, ggplot2::aes(color = species), na.rm = TRUE) +
    ggplot2::scale_linetype_manual(values = c("Day" = "solid", "Night" = "dashed")) +
    ggplot2::scale_y_reverse(limits = rev(y_lim), expand = c(0.01, 0.01)) +
    ggplot2::labs(x = NULL, y = NULL, linetype = "Period", color = NULL) +
    proposal_theme() +
    ggplot2::scale_color_manual(values = c(
      "Large pelagic" = unname(p$my_palette["largePel"]),
      "Mesopelagic" = unname(p$my_palette["mesoPel"]),
      "Midwater predator" = unname(p$my_palette["midwPred"])
    ), drop = FALSE) +
    ggplot2::theme(
      aspect.ratio = 1.55,
      panel.grid.major.y = ggplot2::element_line(color = "grey88", linewidth = 0.25),
      panel.grid.major.x = ggplot2::element_line(color = "grey90", linewidth = 0.2),
      legend.position = "none",
      axis.title.y = ggplot2::element_blank(),
      axis.text.y = ggplot2::element_blank(),
      axis.ticks.y = ggplot2::element_blank()
    )
}

forcing_aligned <- data.frame(
  depth_mid_m = z_depth,
  temp_C = stats::approx(x = profile_nc$depth_mid_m, y = profile_nc$temp_C, xout = z_depth, rule = 2)$y,
  pO2_kPa = stats::approx(x = profile_nc$depth_mid_m, y = profile_nc$pO2_kPa, xout = z_depth, rule = 2)$y
)

force_panel <- ggplot2::ggplot() +
  ggplot2::geom_path(data = forcing_aligned, ggplot2::aes(x = temp_C, y = depth_mid_m),
                     linewidth = 1.0, color = "#BA584D", na.rm = TRUE) +
  ggplot2::geom_path(data = forcing_aligned, ggplot2::aes(x = pO2_kPa, y = depth_mid_m),
                     linewidth = 1.0, color = "#6a51a3", na.rm = TRUE) +
  ggplot2::scale_x_continuous(limits = c(0, 25), breaks = c(0, 5, 10, 15, 20, 25), name = NULL) +
  ggplot2::scale_y_reverse(limits = rev(y_lim), expand = c(0.01, 0.01)) +
  ggplot2::labs(y = "Depth (m)", x = NULL) +
  proposal_theme() +
  ggplot2::theme(axis.title.x = ggplot2::element_blank(), aspect.ratio = 1.55, legend.position = "none")

fish_p1 <- mk_species_panel(species_order[1])
fish_p2 <- mk_species_panel(species_order[2])
fish_p3 <- mk_species_panel(species_order[3])

combined_profiles <- (force_panel | fish_p1 | fish_p2 | fish_p3) +
  patchwork::plot_layout(ncol = 4, widths = c(1, 1, 1, 1), guides = "collect") +
  patchwork::plot_annotation(caption = "") &
  ggplot2::theme(plot.margin = ggplot2::margin(5.5, 5.5, 5.5, 5.5))

# single bottom x label for fish profiles group
depth_plot <- (fish_p1 | fish_p2 | fish_p3) +
  patchwork::plot_layout(ncol = 3, guides = "collect") +
  patchwork::plot_annotation(caption = "")

combined_profiles
```

### 3) Fish-profile-only panel (for export)

```{r fish-only-export-plot, fig.height=6.8}
depth_plot
```


### 4) Composite proposal panel (left: profiles, right: networks)

```{r proposal-composite, fig.height=5.2, warning=FALSE}
network_o2 <- plotNetwork(sim, manual_scale_b = TRUE, ref_b = 50) +
  proposal_theme(base_size = 9) +
  ggplot2::theme(panel.grid.major.y = ggplot2::element_blank(), legend.position = "none") +
  ggplot2::scale_y_continuous(breaks = c(-0.2 * p$bottom, -0.8 * p$bottom), labels = c("shallow", "deep")) +
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 90, vjust = 0.5))

network_no_o2 <- plotNetwork(sim_no_o2, manual_scale_b = TRUE, ref_b = 50) +
  proposal_theme(base_size = 9) +
  ggplot2::theme(panel.grid.major.y = ggplot2::element_blank(), legend.position = "none") +
  ggplot2::scale_y_continuous(breaks = c(-0.2 * p_no_o2$bottom, -0.8 * p_no_o2$bottom), labels = c("shallow", "deep")) +
  ggplot2::theme(axis.text.y = ggplot2::element_text(angle = 90, vjust = 0.5))

left_profiles <- patchwork::wrap_plots(list(force_panel, fish_p1, fish_p2, fish_p3),
  ncol = 4, widths = c(1, 1, 1, 1), guides = "collect"
)
right_networks <- patchwork::wrap_plots(list(network_no_o2, network_o2), ncol = 1, nrow = 2)
proposal_panel <- patchwork::wrap_plots(list(left_profiles, right_networks),
  ncol = 2, widths = c(1, 1)
)
proposal_panel
```
```{r proposal-save, eval=TRUE}
out_dir <- here::here("outputs", "proposal_figures")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

ggplot2::ggsave(file.path(out_dir, "cce_forcing_profiles_temp_o2.png"), forcing_plot,
                width = 6.5, height = 7.8, dpi = 300)

ggplot2::ggsave(file.path(out_dir, "cce_forcing_plus_fish_profiles.png"), combined_profiles,
                width = 6.5, height = 9.4, dpi = 300)

ggplot2::ggsave(file.path(out_dir, "cce_largest_stage_day_night_depth_profiles.png"), depth_plot,
                width = 6.5, height = 9.4, dpi = 300)

ggplot2::ggsave(file.path(out_dir, "cce_trophic_network_o2_vs_no_o2.png"), right_networks,
                width = 6.5, height = 10.0, dpi = 300)

ggplot2::ggsave(file.path(out_dir, "cce_profiles_and_networks_composite.png"), proposal_panel,
                width = 6.5, height = 5.2, dpi = 300)
```
