---
title: "Testing `setupVerticalO2()` from a Single NetCDF Coordinate"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing setupVerticalO2 from NetCDF coordinate}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5)
```

## Purpose

This test vignette is a runnable workflow for validating that a **single profile can be read directly from the NetCDF source** and then passed into `setupVerticalO2()`.

Default coordinate is the same CCE location used in the CSV workflow:

- longitude (180 convention): `-123`
- latitude: `34`

## Load FEISTY functions and configure NetCDF input

```{r setup-local}
# Load FEISTY functions locally through devtools.
devtools::load_all(path = here::here())

nc_hist_path <- here::here("data/profiles_hist_last10y.nc")
stopifnot(file.exists(nc_hist_path))

# Default coordinate matches the CCE row used in CSV-based tests.
site_name <- "CCE"
lon180_requested <- -123
lat_requested <- 34
scenario <- "hist"
```

## Read a single coordinate profile from NetCDF

```{r profile-from-nc}
stopifnot(requireNamespace("ncdf4", quietly = TRUE))

nc <- ncdf4::nc_open(nc_hist_path)
on.exit(ncdf4::nc_close(nc), add = TRUE)

lon <- as.numeric(ncdf4::ncvar_get(nc, "lon"))
lat <- as.numeric(ncdf4::ncvar_get(nc, "lat"))
lev <- as.numeric(ncdf4::ncvar_get(nc, "lev"))

lon360_requested <- ifelse(lon180_requested < 0, lon180_requested + 360, lon180_requested)
ix <- which.min(abs(lon - lon360_requested))
iy <- which.min(abs(lat - lat_requested))

lon360_grid <- lon[ix]
lon180_grid <- ifelse(lon360_grid > 180, lon360_grid - 360, lon360_grid)
lat_grid <- lat[iy]

cat("Requested (lon180, lat):", lon180_requested, lat_requested, "\n")
cat("Selected  (lon180, lat):", lon180_grid, lat_grid, "\n")
cat("Grid indices (ix, iy):", ix, iy, "\n")

temp_C <- as.numeric(read_var_lon_lat_lev(nc, "T_mean", length(lon), length(lat), length(lev))[ix, iy, ])
pO2_kPa <- as.numeric(read_var_lon_lat_lev(nc, "pO2_mean", length(lon), length(lat), length(lev))[ix, iy, ])
zmeso <- as.numeric(read_var_lon_lat_lev(nc, "zmeso_mean", length(lon), length(lat), length(lev))[ix, iy, ])
zmicro <- as.numeric(read_var_lon_lat_lev(nc, "zmicro_mean", length(lon), length(lat), length(lev))[ix, iy, ])
I_day_rel <- as.numeric(read_var_lon_lat_lev(nc, "I_day", length(lon), length(lat), length(lev))[ix, iy, ])
I_night_rel <- as.numeric(read_var_lon_lat_lev(nc, "I_night", length(lon), length(lat), length(lev))[ix, iy, ])

bnds <- layer_bounds_from_midpoints(lev)
profile_nc <- data.frame(
  site = site_name,
  lon180_requested = lon180_requested,
  lat_requested = lat_requested,
  lon360_grid = lon360_grid,
  lon180_grid = lon180_grid,
  lat_grid = lat_grid,
  depth_idx = seq_along(lev),
  depth_mid_m = lev,
  depth_top_m = bnds$top_m,
  depth_bot_m = bnds$bot_m,
  dz_m = bnds$dz_m,
  scenario = scenario,
  temp_C = temp_C,
  pO2_kPa = pO2_kPa,
  zmeso = zmeso,
  zmicro = zmicro,
  I_day_rel = I_day_rel,
  I_night_rel = I_night_rel,
  stringsAsFactors = FALSE
)

# Normalize with the same trimming/interpolation logic used by
# read_vertical_o2_profile(...), so profile= and profile_path= behave alike.
profile_nc <- normalize_vertical_o2_profile(
  profile_nc,
  fill_internal_gaps = FALSE
)

str(profile_nc)
head(profile_nc)
summary(profile_nc$dz_m)
```

## Plot vertical forcing profiles

```{r forcing-profiles, fig.height=8}
op <- par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
plot(profile_nc$temp_C, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "Temperature (C)", ylab = "Depth (m)", main = "Temperature",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$pO2_kPa, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "pO2 (kPa)", ylab = "Depth (m)", main = "Oxygen",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$zmicro, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmicro", ylab = "Depth (m)", main = "Microzoo",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$zmeso, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmeso", ylab = "Depth (m)", main = "Mesozoo",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$I_day_rel, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_day_rel", ylab = "Depth (m)", main = "Light (day)",
     ylim = rev(range(profile_nc$depth_mid_m)))
plot(profile_nc$I_night_rel, profile_nc$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_night_rel", ylab = "Depth (m)", main = "Light (night)",
     ylim = rev(range(profile_nc$depth_mid_m)))
par(op)
```

## Build `setupVerticalO2()` from NetCDF-derived profile object

```{r build-setup}
p <- setupVerticalO2(
  profile = profile_nc,
  nStages = 9,
  K_O = 2,
  h_O = 1,
  delta_pO2_ref = 2,
  Q10_D = 2,
  Q10_S = 1.5,
  K_L = 0.1,
  L_min = 0.5,
  L_max = 1.5,
  light_bottom_to_one = TRUE
)

# Compare to profile_path path behavior: setupVerticalO2() only reads from
# profile_path when profile is NULL.
cat("nrow(profile_nc):", nrow(profile_nc), "\n")
cat("depth range:", min(profile_nc$depth_mid_m), "to", max(profile_nc$depth_mid_m), "m\n")

p$setup
dim(p$depthDay)
dim(p$depthNight)
range(p$glvl)
range(p$phiLightDay)
range(p$phiLightNight)
```

## Run and inspect a short simulation

```{r run-sim}
sim <- simulateFEISTY(p = p, tEnd = 50, USEdll = FALSE)
```
