---
title: "Serial `setupVerticalO2()` site examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Serial setupVerticalO2 site examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5)
```

## Purpose

This vignette runs a small serial comparison across sites in
`data/profiles_sites_hist_vs_ssp585_long.csv`. It intentionally avoids geographic plots,
NetCDF access, and parallel execution. The goal is to show how to split a loaded
profile table into individual profiles and pass each one to `setupVerticalO2()`.

## Load profiles

```{r load-profiles}
devtools::load_all(path = here::here())
source(here::here("scripts", "vertical_o2_profiles.R"))

profile_path <- here::here("data", "profiles_sites_hist_vs_ssp585_long.csv")
profile_all <- read_vertical_o2_profile(profile_path = profile_path, scenario = "hist")
profile_all$key <- paste(profile_all$site, profile_all$scenario, sep = "::")
profiles <- split(profile_all, profile_all$key)

names(profiles)
```

## Run one setup and short simulation per site

```{r serial-run}
run_one_profile <- function(profile_one, nStages = 9, tEnd = 30) {
  profile_one$key <- NULL
  p <- setupVerticalO2(
    profile = profile_one,
    nStages = nStages,
    use_oxygen = TRUE,
    use_light = TRUE
  )
  sim <- simulateFEISTY(p = p, tEnd = tEnd, USEdll = FALSE)
  list(profile = profile_one, p = p, sim = sim)
}

results <- lapply(profiles, run_one_profile)
```

## Summarize final biomass and oxygen limitation

```{r summarize-results}
summarise_result <- function(result, key) {
  sim <- result$sim
  p <- result$p
  final_u <- sim$u[nrow(sim$u), ]
  data.frame(
    key = key,
    site = result$profile$site[1],
    scenario = result$profile$scenario[1],
    total_fish_biomass = sum(final_u[p$ixFish]),
    mean_glvl = mean(p$glvl[p$ixFish]),
    min_pO2_kPa = min(result$profile$pO2_kPa),
    stringsAsFactors = FALSE
  )
}

summary_df <- do.call(rbind, Map(summarise_result, results, names(results)))
summary_df
```

## Plot site summaries

```{r plot-site-summary, fig.height=4.8}
op <- par(mfrow = c(1, 2), mar = c(6, 4, 2, 1))
barplot(summary_df$total_fish_biomass, names.arg = summary_df$site, las = 2,
        ylab = "final fish biomass", main = "Short-run biomass")
barplot(summary_df$mean_glvl, names.arg = summary_df$site, las = 2,
        ylab = "mean glvl", main = "Mean oxygen limitation", ylim = c(0, 1))
par(op)
```

## Plot one example simulation

```{r plot-one-example, fig.height=6}
example_key <- names(results)[1]
plotSimulation(results[[example_key]]$sim)
```
