---
title: "Temperature + Oxygen Regime Maps from NetCDF Forcing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Temperature and oxygen regime maps from NetCDF forcing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 2.6)
```

## Purpose

Build **three global maps** (surface, midwater, bottom) from historical and projected NetCDF forcing files by:

1. averaging projected minus historical **temperature** and **oxygen** differences within each depth regime,
2. classifying each grid cell as:
   - `Warming`
   - `Deoxygenation`
   - `Warming & Deoxygenation`
   - `No regime change`
3. plotting in **Robinson projection** with total panel width 6.5".

## User controls

```{r controls}
nc_hist_path <- here::here("data", "profiles_hist_last10y.nc")
nc_proj_path <- here::here("data", "profiles_proj_last10y.nc")
out_dir <- here::here("outputs", "temp_o2_regime_maps")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Depth breaks in meters.
surface_max_m <- 200
midwater_max_m <- 1000

# Thresholds for classification.
warming_threshold_C <- 0.5
deoxygenation_threshold_kPa <- -0.25

stopifnot(file.exists(nc_hist_path), file.exists(nc_proj_path))
```

## Read NetCDF and align variables

```{r read-nc}
stopifnot(requireNamespace("ncdf4", quietly = TRUE))
stopifnot(requireNamespace("dplyr", quietly = TRUE))

nc_h <- ncdf4::nc_open(nc_hist_path)
on.exit(ncdf4::nc_close(nc_h), add = TRUE)
nc_p <- ncdf4::nc_open(nc_proj_path)
on.exit(ncdf4::nc_close(nc_p), add = TRUE)

lon_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lon"))
lat_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lat"))
lev_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lev"))

lon_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lon"))
lat_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lat"))
lev_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lev"))

stopifnot(identical(lon_h, lon_p), identical(lat_h, lat_p), identical(lev_h, lev_p))

T_hist <- read_var_lon_lat_lev(nc_h, "T_mean", length(lon_h), length(lat_h), length(lev_h))
T_proj <- read_var_lon_lat_lev(nc_p, "T_mean", length(lon_h), length(lat_h), length(lev_h))
O_hist <- read_var_lon_lat_lev(nc_h, "pO2_mean", length(lon_h), length(lat_h), length(lev_h))
O_proj <- read_var_lon_lat_lev(nc_p, "pO2_mean", length(lon_h), length(lat_h), length(lev_h))

dT <- T_proj - T_hist
dO <- O_proj - O_hist

lon180 <- ifelse(lon_h > 180, lon_h - 360, lon_h)
```

## Build depth regimes and summarize mean deltas

```{r summarize-by-regime}
regime_from_depth <- function(depth_m, depth_vec) {
  deepest_mid <- max(depth_vec[is.finite(depth_vec)])

  if (!is.finite(depth_m)) return(NA_character_)
  if (depth_m >= deepest_mid) return("bottom")
  if (depth_m <= surface_max_m) return("surface")
  if (depth_m <= midwater_max_m) return("midwater")
  "bottom"
}

regime_levels <- c("surface", "midwater", "bottom")

cells <- expand.grid(
  ix = seq_along(lon_h),
  iy = seq_along(lat_h),
  iz = seq_along(lev_h),
  KEEP.OUT.ATTRS = FALSE
)

cells$lon <- lon180[cells$ix]
cells$lat <- lat_h[cells$iy]
cells$depth_mid_m <- lev_h[cells$iz]
cells$depth_regime <- vapply(cells$depth_mid_m, regime_from_depth, character(1), depth_vec = lev_h)
cells$dT <- as.numeric(dT[cbind(cells$ix, cells$iy, cells$iz)])
cells$dO <- as.numeric(dO[cbind(cells$ix, cells$iy, cells$iz)])

cells <- cells[is.finite(cells$dT) & is.finite(cells$dO), , drop = FALSE]

delta_regimes <- cells |>
  dplyr::group_by(depth_regime, lon, lat) |>
  dplyr::summarise(
    delta_temp_C = mean(dT, na.rm = TRUE),
    delta_o2_kPa = mean(dO, na.rm = TRUE),
    .groups = "drop"
  )

delta_regimes$depth_regime <- factor(delta_regimes$depth_regime, levels = regime_levels)
```

## Classify regime change

```{r classify}
delta_regimes$regime_class <- with(
  delta_regimes,
  ifelse(
    delta_temp_C >= warming_threshold_C & delta_o2_kPa <= deoxygenation_threshold_kPa,
    "Warming & Deoxygenation",
    ifelse(
      delta_temp_C >= warming_threshold_C,
      "Warming",
      ifelse(delta_o2_kPa <= deoxygenation_threshold_kPa, "Deoxygenation", "No regime change")
    )
  )
)

delta_regimes$regime_class <- factor(
  delta_regimes$regime_class,
  levels = c("Warming", "Deoxygenation", "Warming & Deoxygenation", "No regime change")
)

head(delta_regimes)
```

## Plot 3-panel Robinson maps

```{r map, fig.width=6.5, fig.height=2.8}
stopifnot(requireNamespace("sf", quietly = TRUE))
stopifnot(requireNamespace("rnaturalearth", quietly = TRUE))
stopifnot(requireNamespace("ggplot2", quietly = TRUE))

world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
robin_crs <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

infer_cell_size <- function(x) {
  ux <- sort(unique(x[is.finite(x)]))
  if (length(ux) < 2) return(1)
  d <- min(diff(ux), na.rm = TRUE)
  if (!is.finite(d) || d <= 0) 1 else d
}

class_cols <- c(
  "Warming" = "#E69F00",
  "Deoxygenation" = "#56B4E9",
  "Warming & Deoxygenation" = "#CC79A7",
  "No regime change" = "#D9D9D9"
)

p_regimes <- ggplot2::ggplot(
  delta_regimes,
  ggplot2::aes(x = lon, y = lat, fill = regime_class)
) +
  ggplot2::geom_tile(
    width = infer_cell_size(delta_regimes$lon),
    height = infer_cell_size(delta_regimes$lat)
  ) +
  ggplot2::geom_sf(data = world, fill = "grey20", color = "grey20", linewidth = 0.05, inherit.aes = FALSE) +
  ggplot2::scale_fill_manual(values = class_cols, drop = FALSE) +
  ggplot2::coord_sf(crs = sf::st_crs(robin_crs), default_crs = sf::st_crs(4326), expand = FALSE) +
  ggplot2::facet_wrap(~ depth_regime, nrow = 1) +
  ggplot2::labs(
    title = "Projected - Historical temperature & oxygen regime change",
    subtitle = sprintf(
      "Thresholds: ΔT ≥ %.2f °C, ΔO2 ≤ %.2f kPa",
      warming_threshold_C, deoxygenation_threshold_kPa
    ),
    fill = "Class"
  ) +
  ggplot2::theme_minimal(base_size = 8) +
  ggplot2::theme(
    panel.grid = ggplot2::element_blank(),
    axis.text = ggplot2::element_blank(),
    axis.title = ggplot2::element_blank(),
    strip.text = ggplot2::element_text(face = "bold"),
    legend.position = "bottom"
  )

plot_file <- file.path(out_dir, "temp_o2_regime_maps_robinson_6p5in.png")
ggplot2::ggsave(plot_file, p_regimes, width = 6.5, height = 2.8, dpi = 300)

p_regimes
```
