---
title: "Temperature + Oxygen Regime Maps from NetCDF Forcing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Temperature and oxygen regime maps from NetCDF forcing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 6.5, fig.height = 2.6)
```

## Purpose

Build **three global maps** (surface, midwater, bottom) from historical and projected NetCDF forcing files by:

1. averaging projected minus historical **temperature** and **oxygen** differences within each depth regime,
2. classifying each grid cell as:
   - `Warming`
   - `Deoxygenation`
   - `Warming & Deoxygenation`
   - `No regime change`
3. plotting in **Robinson projection** with total panel width 6.5".

## Load FEISTY functions and configure NetCDF inputs

```{r setup-local}
# Load FEISTY functions locally through devtools (same pattern as other setupVerticalO2 NetCDF vignettes).
devtools::load_all(path = here::here())

nc_hist_path <- here::here("data", "profiles_hist_last10y.nc")
nc_proj_path <- here::here("data", "profiles_ssp585_last10y.nc")
out_dir <- here::here("outputs", "temp_o2_regime_maps")

dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)

# Depth breaks in meters.
surface_max_m <- 200
midwater_max_m <- 1000

# Thresholds for classification.
warming_threshold_C <- 0.5
deoxygenation_threshold_kPa <- -0.25

stopifnot(file.exists(nc_hist_path), file.exists(nc_proj_path))
```

## Read NetCDF and align variables

```{r read-nc}
stopifnot(requireNamespace("ncdf4", quietly = TRUE))
stopifnot(requireNamespace("dplyr", quietly = TRUE))

nc_h <- ncdf4::nc_open(nc_hist_path)
on.exit(ncdf4::nc_close(nc_h), add = TRUE)
nc_p <- ncdf4::nc_open(nc_proj_path)
on.exit(ncdf4::nc_close(nc_p), add = TRUE)

lon_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lon"))
lat_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lat"))
lev_h <- as.numeric(ncdf4::ncvar_get(nc_h, "lev"))

lon_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lon"))
lat_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lat"))
lev_p <- as.numeric(ncdf4::ncvar_get(nc_p, "lev"))

stopifnot(identical(lon_h, lon_p), identical(lat_h, lat_p), identical(lev_h, lev_p))

T_hist <- read_var_lon_lat_lev(nc_h, "T_mean", length(lon_h), length(lat_h), length(lev_h))
T_proj <- read_var_lon_lat_lev(nc_p, "T_mean", length(lon_h), length(lat_h), length(lev_h))
O_hist <- read_var_lon_lat_lev(nc_h, "pO2_mean", length(lon_h), length(lat_h), length(lev_h))
O_proj <- read_var_lon_lat_lev(nc_p, "pO2_mean", length(lon_h), length(lat_h), length(lev_h))

dT <- T_proj - T_hist
dO <- O_proj - O_hist

lon180 <- ifelse(lon_h > 180, lon_h - 360, lon_h)
```

## Build depth regimes and summarize mean deltas

```{r summarize-by-regime}
regime_from_depth <- function(depth_m, depth_vec) {
  deepest_mid <- max(depth_vec[is.finite(depth_vec)])

  if (!is.finite(depth_m)) return(NA_character_)
  if (depth_m >= deepest_mid) return("bottom")
  if (depth_m <= surface_max_m) return("surface")
  if (depth_m <= midwater_max_m) return("midwater")
  "bottom"
}

regime_levels <- c("surface", "midwater", "bottom")

cells <- expand.grid(
  ix = seq_along(lon_h),
  iy = seq_along(lat_h),
  iz = seq_along(lev_h),
  KEEP.OUT.ATTRS = FALSE
)

cells$lon <- lon180[cells$ix]
cells$lat <- lat_h[cells$iy]
cells$depth_mid_m <- lev_h[cells$iz]
cells$depth_regime <- vapply(cells$depth_mid_m, regime_from_depth, character(1), depth_vec = lev_h)
cells$dT <- as.numeric(dT[cbind(cells$ix, cells$iy, cells$iz)])
cells$dO <- as.numeric(dO[cbind(cells$ix, cells$iy, cells$iz)])

cells <- cells[is.finite(cells$dT) & is.finite(cells$dO), , drop = FALSE]

delta_regimes <- cells |>
  dplyr::group_by(depth_regime, lon, lat) |>
  dplyr::summarise(
    delta_temp_C = mean(dT, na.rm = TRUE),
    delta_o2_kPa = mean(dO, na.rm = TRUE),
    .groups = "drop"
  )

delta_regimes$depth_regime <- factor(
  delta_regimes$depth_regime,
  levels = regime_levels,
  labels = c("Surface", "Midwater", "Bottom")
)
```

## Classify regime change

```{r classify}
delta_regimes$regime_class <- with(
  delta_regimes,
  ifelse(
    delta_temp_C >= warming_threshold_C & delta_o2_kPa <= deoxygenation_threshold_kPa,
    "Warming & Deoxygenation",
    ifelse(
      delta_temp_C >= warming_threshold_C,
      "Warming",
      ifelse(delta_o2_kPa <= deoxygenation_threshold_kPa, "Deoxygenation", "No regime change")
    )
  )
)

delta_regimes$regime_class <- factor(
  delta_regimes$regime_class,
  levels = c("Warming", "Deoxygenation", "Warming & Deoxygenation", "No regime change")
)

head(delta_regimes)
```

## Plot 3-panel Robinson maps

```{r map, fig.width=6.5, fig.height=6.8}
stopifnot(requireNamespace("sf", quietly = TRUE))
stopifnot(requireNamespace("terra", quietly = TRUE))
stopifnot(requireNamespace("rnaturalearth", quietly = TRUE))
stopifnot(requireNamespace("ggplot2", quietly = TRUE))
stopifnot(requireNamespace("patchwork", quietly = TRUE))

world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
robin_crs <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"
world_robin <- sf::st_transform(world, crs = sf::st_crs(robin_crs))

class_cols <- c(
  "Warming" = "#C9574A",
  "Deoxygenation" = "#674776",
  "Warming & Deoxygenation" = "#CC79A7",
  "No regime change" = "#D3D694"
)

class_levels <- levels(delta_regimes$regime_class)
delta_regimes$class_code <- as.integer(factor(delta_regimes$regime_class, levels = class_levels))

lon_vals <- sort(unique(delta_regimes$lon))
lat_vals <- sort(unique(delta_regimes$lat))

dlon <- min(diff(lon_vals))
dlat <- min(diff(lat_vals))

make_projected_raster_df <- function(dat_regime) {
  r <- terra::rast(
    ncols = length(lon_vals),
    nrows = length(lat_vals),
    xmin = min(lon_vals) - dlon / 2,
    xmax = max(lon_vals) + dlon / 2,
    ymin = min(lat_vals) - dlat / 2,
    ymax = max(lat_vals) + dlat / 2,
    crs = "EPSG:4326"
  )

  vals <- rep(NA_real_, terra::ncell(r))
  col_ix <- match(dat_regime$lon, lon_vals)
  row_ix <- match(dat_regime$lat, lat_vals)
  cell_ix <- terra::cellFromRowCol(r, row = length(lat_vals) - row_ix + 1, col = col_ix)
  vals[cell_ix] <- dat_regime$class_code
  terra::values(r) <- vals

  r_robin <- terra::project(r, robin_crs, method = "near")
  df <- as.data.frame(r_robin, xy = TRUE, na.rm = FALSE)
  names(df)[3] <- "class_code"
  df$regime_class <- factor(class_levels[df$class_code], levels = class_levels)
  df
}

plot_regime <- function(regime_label) {
  dat <- subset(delta_regimes, depth_regime == regime_label)
  df_r <- make_projected_raster_df(dat)

  ggplot2::ggplot(df_r, ggplot2::aes(x = x, y = y, fill = regime_class)) +
    ggplot2::geom_raster(interpolate = FALSE) +
    ggplot2::geom_sf(data = world_robin, fill = "grey20", color = "grey20", linewidth = 0.05, inherit.aes = FALSE) +
    ggplot2::scale_fill_manual(values = class_cols, drop = FALSE) +
    ggplot2::coord_sf(crs = sf::st_crs(robin_crs), expand = FALSE) +
    ggplot2::labs(title = regime_label, fill = "Class") +
    ggplot2::theme_minimal(base_size = 8) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text = ggplot2::element_blank(),
      axis.title = ggplot2::element_blank(),
      plot.title = ggplot2::element_text(face = "bold", hjust = 0.5),
      legend.position = "bottom",
      legend.direction = "horizontal"
    )
}

p_surface <- plot_regime("Surface")
p_midwater <- plot_regime("Midwater")
p_bottom <- plot_regime("Bottom")

p_regimes <- (
  p_surface + p_midwater
) / (
  patchwork::plot_spacer() + p_bottom + patchwork::plot_spacer()
) +
  patchwork::plot_layout(guides = "collect") &
  ggplot2::theme(legend.position = "bottom", legend.direction = "horizontal")

plot_file_png <- file.path(out_dir, "temp_o2_regime_maps_robinson_6p5in.png")
plot_file_pdf <- file.path(out_dir, "temp_o2_regime_maps_robinson_6p5in.pdf")

ggplot2::ggsave(plot_file_png, p_regimes, width = 6.5, height = 6.8, dpi = 300)
ggplot2::ggsave(plot_file_pdf, p_regimes, width = 6.5, height = 6.8, device = grDevices::cairo_pdf)

p_regimes
```
