---
title: "Testing `setupVerticalO2()` Locally"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing setupVerticalO2 Locally}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5)
```

## Purpose

This test vignette is a runnable workflow for validating `setupVerticalO2()` with:

1. Local loading of FEISTY code and profile data.
2. Vertical profile and habitat/overlap diagnostics.
3. A short simulation run.
4. Food web and simulation summary plots.

## Load FEISTY functions and profile data locally

```{r setup-local}
# Load FEISTY functions locally through devtools.
devtools::load_all(path = here::here())

profile_path <- here::here("data/profiles_sites_hist_vs_ssp585_long.csv")
stopifnot(file.exists(profile_path))

site <- "CCE"
scenario <- "hist"
```

## Read and inspect a profile

```{r profile-load}
prof <- read_vertical_o2_profile(
  profile_path = profile_path,
  site = site,
  scenario = scenario,
  fill_internal_gaps = FALSE
)

str(prof)
head(prof)
summary(prof$dz_m)
```

## Plot vertical forcing profiles

```{r forcing-profiles, fig.height=8}
op <- par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
plot(prof$temp_C, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "Temperature (C)", ylab = "Depth (m)", main = "Temperature")
plot(prof$pO2_kPa, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "pO2 (kPa)", ylab = "Depth (m)", main = "Oxygen")
plot(prof$zmicro, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmicro", ylab = "Depth (m)", main = "Microzoo")
plot(prof$zmeso, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmeso", ylab = "Depth (m)", main = "Mesozoo")
plot(prof$I_day_rel, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_day_rel", ylab = "Depth (m)", main = "Light (day)")
plot(prof$I_night_rel, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_night_rel", ylab = "Depth (m)", main = "Light (night)")
for (i in 1:6) {
  usr <- par("usr")
  # Flip y-axis so surface appears at top
  par(usr = c(usr[1], usr[2], max(prof$depth_mid_m), min(prof$depth_mid_m)))
  box()
}
par(op)
```

## Build `setupVerticalO2()`

```{r build-setup}
p <- setupVerticalO2(
  profile_path = profile_path,
  site = site,
  scenario = scenario,
  nStages = 9,
  K_O = 2,
  h_O = 1,
  delta_pO2_ref = 2,
  Q10_D = 2,
  Q10_S = 1.5,
  K_L = 0.1,
  L_min = 0.5,
  light_bottom_to_one = TRUE
)

p$setup
dim(p$depthDay)
dim(p$depthNight)
range(p$glvl)
range(p$phiLightDay)
range(p$phiLightNight)
```

## Vertical-distribution and oxygen/light diagnostics

```{r diagnostics, fig.height=8}
op <- par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))

matplot(p$depthDay[, p$ixFish], p$z_mid, type = "l", lty = 1,
        xlab = "Depth fraction", ylab = "Depth (m)",
        main = "DepthDay by fish stage")

plot(p$phiLightDay, p$z_mid, type = "l", lwd = 2,
     xlab = "Light scalar", ylab = "Depth (m)",
     main = "phiLightDay / phiLightNight")
lines(p$phiLightNight, p$z_mid, lwd = 2, lty = 2)
legend("bottomright", legend = c("Day", "Night"), lty = c(1,2), bty = "n")

matplot(p$glvl_day[, p$ixFish], p$z_mid, type = "l", lty = 1,
        xlab = "Local glvl", ylab = "Depth (m)",
        main = "glvl_day by fish stage")

plot(p$glvl[p$ixFish], type = "h", lwd = 3,
     xlab = "Fish stage index", ylab = "Effective glvl",
     main = "Stage-effective glvl")

for (k in 1:3) {
  usr <- par("usr")
  par(usr = c(usr[1], usr[2], max(p$z_mid), min(p$z_mid)))
  box()
}
par(op)
```

## Run and inspect a short simulation

```{r run-sim}
sim <- simulateFEISTY(p = p, tEnd = 50)
head(sim)
```

## Food-web structure and rate plots

```{r foodweb-plots, fig.height=6}
plotNetwork(sim)
plotDiet(sim)
plotRates(sim)
```

```{r sim-plots, fig.height=6}
plotSimulation(sim)
plotBiomasstime(sim)
plotYieldtime(sim)
```

## Notes for global loops

For global workflows, you can preload the CSV once, split by site/scenario, and call
`setupVerticalO2(profile = one_site_profile, ...)` to avoid repeated file I/O.
