---
title: "Testing `setupVerticalO2()` Locally"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Testing setupVerticalO2 Locally}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5)
```

## Purpose

This test vignette is a runnable workflow for validating `setupVerticalO2()` with:

1. Local loading of FEISTY code and profile data.
2. Vertical profile and habitat/overlap diagnostics.
3. A short simulation run.
4. Food web and simulation summary plots.

## Load FEISTY functions and profile data locally

```{r setup-local}
# Load FEISTY functions locally through devtools.
devtools::load_all(path = here::here())

profile_path <- here::here("data/profiles_sites_hist_vs_ssp585_long.csv")
stopifnot(file.exists(profile_path))

site <- "CCE"
scenario <- "hist"
```

## Read and inspect a profile

```{r profile-load}
prof <- read_vertical_o2_profile(
  profile_path = profile_path,
  site = site,
  scenario = scenario,
  fill_internal_gaps = FALSE
)

str(prof)
head(prof)
summary(prof$dz_m)
```

```{r optional-constant-o2, eval=FALSE}
# Optional sensitivity toggle (comment-out-able):
# Replace the selected profile oxygen with a constant 30 kPa at all depths.
# This is useful to isolate oxygen effects on g (growth) vs other mechanisms.
prof_30kpa <- prof
prof_30kpa$pO2_kPa <- 30
```

## Plot vertical forcing profiles

```{r forcing-profiles, fig.height=8}
op <- par(mfrow = c(2, 3), mar = c(4, 4, 2, 1))
plot(prof$temp_C, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "Temperature (C)", ylab = "Depth (m)", main = "Temperature",
     ylim = rev(range(prof$depth_mid_m)))
plot(prof$pO2_kPa, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "pO2 (kPa)", ylab = "Depth (m)", main = "Oxygen",
     ylim = rev(range(prof$depth_mid_m)))
plot(prof$zmicro, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmicro", ylab = "Depth (m)", main = "Microzoo",
     ylim = rev(range(prof$depth_mid_m)))
plot(prof$zmeso, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "zmeso", ylab = "Depth (m)", main = "Mesozoo",
     ylim = rev(range(prof$depth_mid_m)))
plot(prof$I_day_rel, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_day_rel", ylab = "Depth (m)", main = "Light (day)",
     ylim = rev(range(prof$depth_mid_m)))
plot(prof$I_night_rel, prof$depth_mid_m, type = "l", lwd = 2,
     xlab = "I_night_rel", ylab = "Depth (m)", main = "Light (night)",
     ylim = rev(range(prof$depth_mid_m)))
par(op)
```

## Build `setupVerticalO2()`

```{r build-setup}
p <- setupVerticalO2(
  profile_path = profile_path,
  site = site,
  scenario = scenario,
  nStages = 9,
  K_O = 2,
  h_O = 1,
  delta_pO2_ref = 2,
  Q10_D = 2,
  Q10_S = 1.5,
  K_L = 0.1,
  L_min = 0.5,
  L_max = 1.5,
  light_bottom_to_one = TRUE
)

# Optional constant-O2 run (uncomment to use):
# p <- setupVerticalO2(
#   profile = prof_30kpa,
#   nStages = 9,
#   K_O = 2,
#   h_O = 1,
#   delta_pO2_ref = 2,
#   Q10_D = 2,
#   Q10_S = 1.5,
#   K_L = 0.1,
#   L_min = 0.5,
#   L_max = 1.5,
#   light_bottom_to_one = TRUE
# )

p$setup
dim(p$depthDay)
dim(p$depthNight)
range(p$glvl)
range(p$phiLightDay)
range(p$phiLightNight)
```

## Vertical-distribution and oxygen/light diagnostics

```{r diagnostics, fig.height=8}
op <- par(mfrow = c(2, 2), mar = c(4, 4, 2, 1))

# Stage colors by fish group
group_cols <- unlist(mapply(function(ix, nm) rep(p$my_palette[nm], length(ix)),
                            p$ix, p$groupnames[p$ixFish], SIMPLIFY = FALSE))
fish_stage_labels <- p$stagenames[p$ixFish]

matplot(p$depthDay[, p$ixFish], p$z_mid, type = "l", lty = 1, col = group_cols,
        xlab = "Depth fraction", ylab = "Depth (m)",
        main = "DepthDay by fish stage", ylim = rev(range(p$z_mid)))
legend("bottomright", legend = p$groupnames[p$ixFish], col = p$my_palette[p$groupnames[p$ixFish]],
       lty = 1, cex = 0.7, bty = "n")

plot(p$phiLightDay, p$z_mid, type = "l", lwd = 2,
     xlab = "Light scalar", ylab = "Depth (m)",
     main = "phiLightDay / phiLightNight", ylim = rev(range(p$z_mid)))
lines(p$phiLightNight, p$z_mid, lwd = 2, lty = 2)
legend("bottomright", legend = c("Day", "Night"), lty = c(1,2), bty = "n")

matplot(p$depthNight[, p$ixFish], p$z_mid, type = "l", lty = 1, col = group_cols,
        xlab = "Depth fraction", ylab = "Depth (m)",
        main = "DepthNight by fish stage", ylim = rev(range(p$z_mid)))

par(new = FALSE)
plot(p$glvl[p$ixFish], type = "h", lwd = 3, col = group_cols,
     xlab = "Fish stage index", ylab = "Effective glvl",
     main = "Stage-effective glvl")
points(p$glvl[p$ixFish], pch = 16, col = group_cols)
axis(1, at = seq_along(p$ixFish), labels = fish_stage_labels, las = 2, cex.axis = 0.55)
abline(h = c(0.25, 0.5, 0.75, 1.0), lty = 3, col = "grey80")
legend("bottomleft", legend = p$groupnames[p$ixFish], col = p$my_palette[p$groupnames[p$ixFish]],
       lty = 1, pch = 16, cex = 0.7, bty = "n")

par(op)
```

## Oxygen-local diagnostics by depth (day vs night)

```{r glvl-local-depth, fig.height=6}
op <- par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
matplot(p$glvl_day[, p$ixFish], p$z_mid, type = "l", lty = 1, col = group_cols,
        xlab = "Local glvl", ylab = "Depth (m)",
        main = "glvl_day by fish stage", ylim = rev(range(p$z_mid)))
matplot(p$glvl_night[, p$ixFish], p$z_mid, type = "l", lty = 1, col = group_cols,
        xlab = "Local glvl", ylab = "Depth (m)",
        main = "glvl_night by fish stage", ylim = rev(range(p$z_mid)))
par(op)
```

## Run and inspect a short simulation

```{r run-sim}
sim <- simulateFEISTY(p = p, tEnd = 50)
head(sim)
```

## Food-web structure and rate plots

```{r foodweb-plots, fig.height=6}
plotNetwork(sim)
plotDiet(sim)
plotRates(sim)
```

```{r sim-plots, fig.height=6}
plotSimulation(sim)
plotBiomasstime(sim)
plotYieldtime(sim)
```

## Notes for global loops

For global workflows, you can preload the CSV once, split by site/scenario, and call
`setupVerticalO2(profile = one_site_profile, ...)` to avoid repeated file I/O.
