---
title: "Parallel `setupVerticalO2()` site comparison maps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Parallel setupVerticalO2 site comparison maps}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 9, fig.height = 5)
```

## Purpose

This vignette runs `setupVerticalO2()` for any number of profile locations,
compares simulations **with oxygen limitation** against a **no-oxygen-limitation**
control (`glvl = 1` everywhere), saves compact biomass summaries, and maps the
results in a Robinson projection with black landmasses.

The workflow is designed for large profile CSV files:

1. Read the profile table once.
2. Split profiles by `(site, scenario)`.
3. Run each site in parallel.
4. Save only compact functional-group biomass summaries.
5. Plot total fish community biomass and oxygen-effect difference maps.

## User controls

Edit this chunk for your run. `locations` can be any number of sites. Use
`locations <- NULL` to run every site/scenario combination available in the
profile file.

```{r controls}
devtools::load_all(path = here::here())

profile_path <- here::here("data/profiles_sites_hist_vs_ssp585_long.csv")
scenario_default <- "hist"

# Any number of site names, or NULL for all sites in the profile file.
site_names <- c("CCE", "HOTstationALOHA")

# Optional explicit location table. If NULL, it is built from site_names and
# scenario_default; if site_names is also NULL, all profile site/scenario pairs
# are used.
locations <- NULL

# Simulation controls. USEdll = FALSE avoids requiring compiled Fortran and is
# the default behavior in this branch.
nStages <- 9
tEnd <- 50
tStep <- 1
n_cores <- max(1, parallel::detectCores() - 1)

out_dir <- here::here("outputs", "setupVerticalO2_parallel_maps")
dir.create(out_dir, recursive = TRUE, showWarnings = FALSE)
```

## Optional packages for mapping

Mapping requires `sf`, `rnaturalearth`, and `rnaturalearthdata`. The simulation
and CSV outputs can still run without them; map chunks are skipped with an
informative message if they are unavailable.

```{r package-check}
map_pkgs <- c("sf", "rnaturalearth", "rnaturalearthdata")
missing_map_pkgs <- map_pkgs[!vapply(map_pkgs, requireNamespace, logical(1), quietly = TRUE)]
can_map <- length(missing_map_pkgs) == 0
if (!can_map) {
  message("Install these packages for maps: ", paste(missing_map_pkgs, collapse = ", "))
}
```

## Load and split the profile table once

For very large CSVs, this chunk uses `data.table::fread()` when available and
falls back to `utils::read.csv()`.

```{r load-profiles}
stopifnot(file.exists(profile_path))

read_profile_table <- function(path) {
  if (requireNamespace("data.table", quietly = TRUE)) {
    as.data.frame(data.table::fread(path))
  } else {
    utils::read.csv(path, stringsAsFactors = FALSE)
  }
}

profile_all <- read_profile_table(profile_path)

if (is.null(locations)) {
  if (is.null(site_names)) {
    locations <- unique(profile_all[, c("site", "scenario")])
  } else {
    locations <- data.frame(
      site = site_names,
      scenario = scenario_default,
      stringsAsFactors = FALSE
    )
  }
}

locations$key <- paste(locations$site, locations$scenario, sep = "::")
profile_all$key <- paste(profile_all$site, profile_all$scenario, sep = "::")
profile_list <- split(profile_all[profile_all$key %in% locations$key, ], profile_all$key[profile_all$key %in% locations$key])

missing_keys <- setdiff(locations$key, names(profile_list))
if (length(missing_keys) > 0) {
  stop("Requested site/scenario combinations not found in profile file: ", paste(missing_keys, collapse = ", "))
}

locations
```

## Worker function

Each worker runs two matched simulations for one profile:

- `use_oxygen = TRUE`: profile oxygen affects `glvl`.
- `use_oxygen = FALSE`: no oxygen limitation (`glvl = 1` everywhere), while the
  same physical/biological profile is otherwise retained.

The saved output is intentionally compact: final functional-group biomasses and
total fish community biomass.

```{r worker}
summarise_final_biomass <- function(sim, condition) {
  p <- sim$p
  final_group <- as.numeric(sim$totBiomass[sim$nTime, ])
  names(final_group) <- p$groupnames[-seq_len(p$nResources)]

  data.frame(
    condition = condition,
    group = names(final_group),
    biomass = final_group,
    stringsAsFactors = FALSE
  )
}

run_one_location <- function(profile_one, key) {
  profile_one <- profile_one[order(profile_one$depth_idx, profile_one$depth_mid_m), ]
  profile_one$key <- NULL

  coord <- profile_one[stats::complete.cases(profile_one[, c("lon180_grid", "lat_grid")]), ][1, ]
  site <- as.character(profile_one$site[1])
  scenario <- as.character(profile_one$scenario[1])

  build_p <- function(use_oxygen) {
    setupVerticalO2(
      profile = profile_one,
      nStages = nStages,
      use_oxygen = use_oxygen,
      K_O = 2,
      h_O = 1,
      delta_pO2_ref = 2,
      Q10_D = 2,
      Q10_S = 1.5,
      K_L = 0.1,
      L_min = 0.5,
      L_max = 1.5,
      light_bottom_to_one = TRUE
    )
  }

  p_o2 <- build_p(TRUE)
  sim_o2 <- simulateFEISTY(p = p_o2, tEnd = tEnd, tStep = tStep, USEdll = FALSE)

  p_no_o2 <- build_p(FALSE)
  sim_no_o2 <- simulateFEISTY(p = p_no_o2, tEnd = tEnd, tStep = tStep, USEdll = FALSE)

  biomass <- rbind(
    summarise_final_biomass(sim_o2, "oxygen_profile"),
    summarise_final_biomass(sim_no_o2, "no_oxygen_limitation")
  )

  biomass$site <- site
  biomass$scenario <- scenario
  biomass$lon <- coord$lon180_grid
  biomass$lat <- coord$lat_grid
  biomass$key <- key

  total <- aggregate(biomass ~ condition + site + scenario + lon + lat + key, biomass, sum)
  total$group <- "total_fish"
  biomass <- rbind(biomass, total[, names(biomass)])

  wide <- reshape(
    biomass[, c("site", "scenario", "lon", "lat", "key", "group", "condition", "biomass")],
    idvar = c("site", "scenario", "lon", "lat", "key", "group"),
    timevar = "condition",
    direction = "wide"
  )
  names(wide) <- sub("^biomass\\.", "biomass_", names(wide))
  wide$biomass_difference <- wide$biomass_oxygen_profile - wide$biomass_no_oxygen_limitation
  wide$biomass_ratio <- wide$biomass_oxygen_profile / wide$biomass_no_oxygen_limitation

  list(biomass = biomass, difference = wide)
}
```

## Run sites in parallel and save outputs

The code uses base R `parallel` so no extra parallel package is required. On
Unix-like systems it uses forking; on Windows it uses a PSOCK cluster.

```{r run-parallel}
keys <- names(profile_list)

if (.Platform$OS.type == "windows" && n_cores > 1) {
  cl <- parallel::makeCluster(n_cores)
  on.exit(parallel::stopCluster(cl), add = TRUE)
  parallel::clusterExport(
    cl,
    varlist = c("profile_list", "run_one_location", "summarise_final_biomass", "nStages", "tEnd", "tStep"),
    envir = environment()
  )
  parallel::clusterEvalQ(cl, devtools::load_all(path = here::here()))
  results <- parallel::parLapply(cl, keys, function(k) run_one_location(profile_list[[k]], k))
} else {
  results <- parallel::mclapply(
    keys,
    function(k) run_one_location(profile_list[[k]], k),
    mc.cores = min(n_cores, length(keys))
  )
}

biomass_summary <- do.call(rbind, lapply(results, `[[`, "biomass"))
biomass_difference <- do.call(rbind, lapply(results, `[[`, "difference"))

utils::write.csv(biomass_summary, file.path(out_dir, "functional_group_biomass_summary.csv"), row.names = FALSE)
utils::write.csv(biomass_difference, file.path(out_dir, "functional_group_biomass_differences.csv"), row.names = FALSE)
saveRDS(results, file.path(out_dir, "setupVerticalO2_parallel_results.rds"))

head(biomass_difference)
```

## Map helpers

```{r map-helpers, eval=can_map}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
robin_crs <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

plot_points_robin <- function(data, value, title, legend_title, palette = ggplot2::scale_color_viridis_c()) {
  pts <- sf::st_as_sf(data, coords = c("lon", "lat"), crs = 4326, remove = FALSE)
  ggplot2::ggplot() +
    ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1) +
    ggplot2::geom_sf(data = pts, ggplot2::aes(color = .data[[value]]), size = 2.5) +
    palette +
    ggplot2::coord_sf(crs = robin_crs) +
    ggplot2::labs(title = title, color = legend_title) +
    ggplot2::theme_minimal() +
    ggplot2::theme(panel.grid.major = ggplot2::element_line(color = "grey85", linewidth = 0.2))
}
```

## Total fish community biomass maps

```{r total-biomass-map, fig.height=5.5, eval=can_map}
total_biomass <- subset(biomass_summary, group == "total_fish")
p_total_biomass <- plot_points_robin(
  subset(total_biomass, condition == "oxygen_profile"),
  value = "biomass",
  title = "Total fish biomass with profile oxygen",
  legend_title = "g m^-2"
)
ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_oxygen_profile.png"), p_total_biomass, width = 10, height = 6, dpi = 300)
p_total_biomass
```

```{r total-difference-map, fig.height=5.5, eval=can_map}
total_diff <- subset(biomass_difference, group == "total_fish")
p_total_diff <- plot_points_robin(
  total_diff,
  value = "biomass_difference",
  title = "Total fish biomass difference: oxygen profile - no oxygen limitation",
  legend_title = "difference (g m^-2)",
  palette = ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)
)
ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_difference.png"), p_total_diff, width = 10, height = 6, dpi = 300)
p_total_diff
```

## Functional-group biomass difference maps

```{r group-difference-maps, fig.height=8, eval=can_map}
group_diff <- subset(biomass_difference, group != "total_fish")
group_diff_sf <- sf::st_as_sf(group_diff, coords = c("lon", "lat"), crs = 4326, remove = FALSE)

p_group_diff <- ggplot2::ggplot() +
  ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1) +
  ggplot2::geom_sf(data = group_diff_sf, ggplot2::aes(color = biomass_difference), size = 2.2) +
  ggplot2::scale_color_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  ggplot2::coord_sf(crs = robin_crs) +
  ggplot2::facet_wrap(~ group) +
  ggplot2::labs(
    title = "Functional group biomass difference: oxygen profile - no oxygen limitation",
    color = "difference (g m^-2)"
  ) +
  ggplot2::theme_minimal() +
  ggplot2::theme(panel.grid.major = ggplot2::element_line(color = "grey85", linewidth = 0.2))

ggplot2::ggsave(file.path(out_dir, "map_functional_group_biomass_differences.png"), p_group_diff, width = 12, height = 8, dpi = 300)
p_group_diff
```

## Notes

- For global runs, set `site_names <- NULL` and increase `n_cores` as needed.
- Outputs are saved in `out_dir`, so the full simulation objects do not need to
  be kept in memory after mapping.
- The no-oxygen control is intentionally `use_oxygen = FALSE`; do not replace
  the observed oxygen profile with an artificial high-pO2 column for this
  comparison.
