---
title: "Baseline community contrasts with `setupVertical()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Baseline setupVertical community contrasts}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 8, fig.height = 5)
```

## Purpose

This vignette provides a baseline reference for how FEISTY communities diverge under the classic `setupVertical()` forcing assumptions.
Use this as a "sanity anchor" when evaluating whether `setupVerticalO2()` is producing enough between-site differentiation.

## Load package locally

```{r setup}
devtools::load_all(path = here::here())
```

## Create contrasting baseline environments

These are not profile-driven; they are controlled contrasts in depth, photic depth, production, and temperature.

```{r configs}
cfg <- data.frame(
  site = c("ShelfWarmHighProd", "SlopeMidProd", "DeepCoolLowProd"),
  depth = c(150, 1200, 3500),
  photic = c(60, 120, 180),
  szprod = c(140, 90, 55),
  lzprod = c(120, 80, 45),
  dfpho = c(500, 320, 180),
  Tp = c(18, 12, 7),
  Tm = c(14, 8, 4),
  Tb = c(10, 5, 2)
)
cfg
```

## Build setups and run simulations

```{r run}
mods <- lapply(seq_len(nrow(cfg)), function(i) {
  setupVertical(
    szprod = cfg$szprod[i],
    lzprod = cfg$lzprod[i],
    dfpho = cfg$dfpho[i],
    depth = cfg$depth[i],
    photic = cfg$photic[i],
    Tp = cfg$Tp[i],
    Tm = cfg$Tm[i],
    Tb = cfg$Tb[i],
    nStages = 9
  )
})
names(mods) <- cfg$site

sims <- lapply(mods, function(p) simulateFEISTY(p = p, tEnd = 120))
```

## Compare biomass time series

```{r biomass}
plots <- lapply(sims, plotBiomasstime)
plots[[1]]
plots[[2]]
plots[[3]]
```

## Compare end-state group biomass

```{r endstate}
extract_end <- function(sim) {
  ts <- getTimeseries(sim)
  te <- max(ts$t)
  sub <- ts[ts$t == te, c("group", "bio")]
  aggregate(bio ~ group, data = sub, FUN = sum)
}

end_list <- Map(function(sim, nm) {
  x <- extract_end(sim)
  x$site <- nm
  x
}, sims, names(sims))

end_df <- do.call(rbind, end_list)
end_df
```

```{r endstate-plot, fig.height=6}
library(ggplot2)

ggplot(end_df, aes(x = group, y = bio, fill = site)) +
  geom_col(position = "dodge") +
  scale_y_log10() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ylab("End biomass (g m-2, log scale)") +
  xlab("Group")
```

## Compare food-web structures

```{r networks, fig.height=6}
plotNetwork(sims[[1]])
plotNetwork(sims[[2]])
plotNetwork(sims[[3]])
```

## Interpretation

If these controlled `setupVertical()` environments produce visibly different end communities,
then limited differentiation under `setupVerticalO2()` likely points to conservative forcing translation
(e.g., overlap geometry, resource diel redistribution, or energy-routing constraints) rather than plotting artifacts.
