---
title: "Plot saved `setupVerticalO2()` map outputs (no rerun)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plot saved setupVerticalO2 map outputs without rerunning parallel analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 9, fig.height = 5)
```

## Purpose

This vignette **only reads saved CSV outputs** from
`setupVerticalO2_parallel_maps.Rmd` and generates the maps, so you do not need
to rerun the full parallel simulation.

It also fixes common plotting failures by:

1. coercing coordinate/value columns to numeric,
2. dropping non-finite rows,
3. aggregating duplicate `(lon, lat, group, condition)` cells,
4. rebuilding `biomass_difference` if needed from `biomass_summary`.

## User controls

```{r controls}
out_dir <- here::here("outputs", "setupVerticalO2_parallel_maps")
summary_file <- file.path(out_dir, "functional_group_biomass_summary.csv")
diff_file <- file.path(out_dir, "functional_group_biomass_differences.csv")

stopifnot(file.exists(summary_file))
```

```{r package-check}
map_pkgs <- c("sf", "rnaturalearth", "rnaturalearthdata", "ggplot2")
missing_map_pkgs <- map_pkgs[!vapply(map_pkgs, requireNamespace, logical(1), quietly = TRUE)]
can_map <- length(missing_map_pkgs) == 0
if (!can_map) {
  message("Install these packages for maps: ", paste(missing_map_pkgs, collapse = ", "))
}
```

## Load and repair saved outputs

```{r load-and-repair}
biomass_summary <- utils::read.csv(summary_file, stringsAsFactors = FALSE)

required_summary_cols <- c("condition", "group", "biomass", "site", "scenario", "lon", "lat", "key")
missing_summary_cols <- setdiff(required_summary_cols, names(biomass_summary))
if (length(missing_summary_cols) > 0) {
  stop("Summary file is missing required columns: ", paste(missing_summary_cols, collapse = ", "))
}

# Coerce columns that frequently break geom_tile when read from CSV
biomass_summary$lon <- suppressWarnings(as.numeric(biomass_summary$lon))
biomass_summary$lat <- suppressWarnings(as.numeric(biomass_summary$lat))
biomass_summary$biomass <- suppressWarnings(as.numeric(biomass_summary$biomass))

biomass_summary <- biomass_summary[
  is.finite(biomass_summary$lon) &
    is.finite(biomass_summary$lat) &
    is.finite(biomass_summary$biomass),
  , drop = FALSE
]

# Aggregate duplicate cells (e.g., multiple scenarios per location)
biomass_summary <- stats::aggregate(
  biomass ~ condition + group + lon + lat,
  data = biomass_summary,
  FUN = mean
)

if (nrow(biomass_summary) == 0) {
  stop("No valid rows remain in biomass_summary after numeric coercion/cleaning.")
}

if (file.exists(diff_file)) {
  biomass_difference <- utils::read.csv(diff_file, stringsAsFactors = FALSE)

  required_diff_cols <- c("group", "lon", "lat", "biomass_difference")
  missing_diff_cols <- setdiff(required_diff_cols, names(biomass_difference))

  if (length(missing_diff_cols) > 0) {
    message("Difference file exists but is missing required columns; rebuilding from summary.")
    biomass_difference <- NULL
  } else {
    biomass_difference$lon <- suppressWarnings(as.numeric(biomass_difference$lon))
    biomass_difference$lat <- suppressWarnings(as.numeric(biomass_difference$lat))
    biomass_difference$biomass_difference <- suppressWarnings(as.numeric(biomass_difference$biomass_difference))

    biomass_difference <- biomass_difference[
      is.finite(biomass_difference$lon) &
        is.finite(biomass_difference$lat) &
        is.finite(biomass_difference$biomass_difference),
      , drop = FALSE
    ]

    biomass_difference <- stats::aggregate(
      biomass_difference ~ group + lon + lat,
      data = biomass_difference,
      FUN = mean
    )
  }
} else {
  biomass_difference <- NULL
}

if (is.null(biomass_difference)) {
  wide <- reshape(
    biomass_summary,
    idvar = c("group", "lon", "lat"),
    timevar = "condition",
    direction = "wide"
  )
  names(wide) <- sub("^biomass\\.", "biomass_", names(wide))

  needed <- c("biomass_oxygen_profile", "biomass_no_oxygen_limitation")
  if (!all(needed %in% names(wide))) {
    stop("Cannot rebuild differences: expected columns ", paste(needed, collapse = ", "), " in reshaped summary.")
  }

  wide$biomass_difference <- wide$biomass_oxygen_profile - wide$biomass_no_oxygen_limitation
  biomass_difference <- wide[, c("group", "lon", "lat", "biomass_difference")]
}

head(biomass_summary)
head(biomass_difference)
```

## Map helpers

```{r map-helpers, eval=can_map}
world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
robin_crs <- "+proj=robin +lon_0=0 +datum=WGS84 +units=m +no_defs"

infer_cell_size <- function(x) {
  ux <- sort(unique(x[is.finite(x)]))
  if (length(ux) < 2) return(1)
  d <- min(diff(ux), na.rm = TRUE)
  if (!is.finite(d) || d <= 0) 1 else d
}

plot_raster_robin <- function(data, value, title, legend_title, palette = ggplot2::scale_fill_viridis_c()) {
  dx <- infer_cell_size(data$lon)
  dy <- infer_cell_size(data$lat)

  ggplot2::ggplot(data, ggplot2::aes(x = lon, y = lat, fill = .data[[value]])) +
    ggplot2::geom_tile(width = dx, height = dy) +
    ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1, inherit.aes = FALSE) +
    palette +
    ggplot2::coord_sf(crs = robin_crs, expand = FALSE) +
    ggplot2::labs(title = title, fill = legend_title) +
    ggplot2::theme_minimal()
}
```

## Total fish community biomass maps

```{r total-biomass-map, fig.height=5.5, eval=can_map}
total_biomass <- subset(biomass_summary, group == "total_fish" & condition == "oxygen_profile")

p_total_biomass <- plot_raster_robin(
  total_biomass,
  value = "biomass",
  title = "Total fish biomass with profile O2",
  legend_title = "g m-2"
)

ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_oxygen_profile_from_saved.png"), p_total_biomass, width = 10, height = 6, dpi = 300)
p_total_biomass
```

```{r total-difference-map, fig.height=5.5, eval=can_map}
total_diff <- subset(biomass_difference, group == "total_fish")

p_total_diff <- plot_raster_robin(
  total_diff,
  value = "biomass_difference",
  title = "Total fish biomass difference: profile O2 - no oxygen limitation",
  legend_title = "Difference (g m-2)",
  palette = ggplot2::scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0)
)

ggplot2::ggsave(file.path(out_dir, "map_total_fish_biomass_difference_from_saved.png"), p_total_diff, width = 10, height = 6, dpi = 300)
p_total_diff
```

## Functional-group biomass difference maps

```{r group-difference-maps, fig.height=8, eval=can_map}
group_diff <- subset(biomass_difference, group != "total_fish")

p_group_diff <- ggplot2::ggplot(group_diff, ggplot2::aes(x = lon, y = lat, fill = biomass_difference)) +
  ggplot2::geom_tile(width = infer_cell_size(group_diff$lon), height = infer_cell_size(group_diff$lat)) +
  ggplot2::geom_sf(data = world, fill = "black", color = "black", linewidth = 0.1, inherit.aes = FALSE) +
  ggplot2::scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  ggplot2::coord_sf(crs = robin_crs, expand = FALSE) +
  ggplot2::facet_wrap(~ group) +
  ggplot2::labs(
    title = "Functional-group biomass difference: profile O2 - no oxygen limitation",
    fill = "Difference (g m-2)"
  ) +
  ggplot2::theme_minimal()

ggplot2::ggsave(file.path(out_dir, "map_functional_group_biomass_differences_from_saved.png"), p_group_diff, width = 12, height = 8, dpi = 300)
p_group_diff
```
