---
title: "FEISTY Vertical Profile Day/Night Analysis Example"
output: html_document
---

```{r setup, message=FALSE, warning=FALSE}
library(devtools)
library(here)
devtools::load_all(here::here())
#library(FEISTY)
```

## 1) Load profile forcing

This example uses the profile file provided for the dynamic vertical workflow:

```{r load-profiles}
profile_path <- here("data", "profiles_sites_hist_vs_ssp585_long.csv")
stopifnot(file.exists(profile_path))

raw <- read.csv(profile_path, stringsAsFactors = FALSE)
site <- unique(raw$site)[1]
scenario <- unique(raw$scenario)[1]

profiles <- read_feisty_vertical_profiles(
  path = profile_path,
  site = site,
  scenario = scenario
)

head(profiles)
```

## 2) Define group-size descriptors and run vertical analysis

`analyze_feisty_vertical_profiles()` produces day/night suitability and occupancy diagnostics across the arbitrary depth grid `z = 1..Z`.

```{r run-analysis}
groups <- c("small_pelagic", "mesopelagic", "large_pelagic", "midwater_predator", "demersal")
is_larval <- rep(FALSE, length(groups))
w_q <- c(1, 10, 100, 50, 20)  # representative masses (g)

ana <- analyze_feisty_vertical_profiles(
  path = profile_path,
  site = site,
  scenario = scenario,
  groups = groups,
  is_larval = is_larval,
  w_q = w_q
)

str(list(
  P_day = dim(ana$P_day),
  P_night = dim(ana$P_night),
  Suit_day = dim(ana$Suit_day),
  Suit_night = dim(ana$Suit_night)
))
```

## 3) Inspect day/night profile diagnostics

```{r plot-input-light, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "input_light")
```

```{r plot-input-o2, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "input_o2")
```

```{r plot-suit-day, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "suitability_day")
```

```{r plot-suit-night, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "suitability_night")
```

## 4) Inspect predicted day/night occupancy distributions

```{r plot-p-day, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "P_day")
```

```{r plot-p-night, fig.width=6, fig.height=5}
plot_feisty_vertical_analysis(ana, variable = "P_night")
```

## 5) Print depth-moment diagnostics

```{r moments}
out <- data.frame(
  group = groups,
  z_mean_day = ana$z_mean_day,
  z_var_day = ana$z_var_day,
  z_mean_night = ana$z_mean_night,
  z_var_night = ana$z_var_night
)
out
```

## Notes

- This vignette demonstrates the current vertical analysis wrapper workflow built around profile-derived forcing and diagnostics.
- It does not yet replace the full FEISTY core simulation loop; instead it provides day/night suitability and occupancy diagnostics needed for the dynamic vertical setup.
